PyQBench: a Python library for benchmarking gate-based quantum computers

### 2023-03-22 txt/manuscript.tex
Add missing article

\subsection{Using user-defined measurement with \texttt{qbench} package}
In this example, we will demonstrate how \texttt{qbench} package can be used with user--defined
measurement. For this purpose, we will use $U = H$ (the Hadamard gate). The detailed calculations
that lead to the particular form of the discriminator and final measurements can be found in
\ref{app:hadamard}.
\noindent
The explicit formula for discriminator in this example reads:
\begin{equation}
\ket{\psi_0} = \frac{1}{\sqrt{2}} (\ket{00} + \ket{11}),
\end{equation}
with final measurements being equal to
\begin{equation}
V_0 =
\left(\begin{array}{cc} \alpha & -\beta\\ \beta & \alpha \end{array}\right),
\end{equation}
and \begin{equation}
V_1 =
\left(\begin{array}{cc} -\beta & \alpha \\ \alpha & \beta \end{array}\right),
\end{equation}
where \begin{equation}
\alpha = \frac{\sqrt{2 - \sqrt{2}}}{2} = \cos\left( \frac{3}{8} \pi \right),
\end{equation}
\begin{equation}
\beta  = \frac{\sqrt{2  + \sqrt{2}}}{2} = \sin\left( \frac{3}{8} \pi \right).
\end{equation}
To use the above benchmarking scheme in PyQBench, we first need to construct circuits that can be executed by actual hardware. To this end, we need to represent each of the unitaries as a sequence of standard gates, keeping in mind that quantum circuits start execution from  $\ket{00}$ state. The circuit taking $\ket{00}$ to the Bell state $\ket{\psi_0}$  comprises the Hadamard gate followed by CNOT gate on both qubits (see Fig.~\ref{fig:discriminator}).
\begin{figure}[h!]
	\centering
	\includegraphics[scale=1.7]{pics/discriminator}
	\caption{Decomposition of the Bell state $\ket{\psi_0}$. }
	\label{fig:discriminator}
\end{figure}
For $V_0$ and $V_1$ observe that $V_0 = \operatorname{RY}\left(\frac{3}{4} \pi \right) $, where $\operatorname{RY}$ is rotation gate around the $Y$ axis defined by
\begin{equation}
\operatorname{RY}(\theta) =
\left(\begin{array}{cc} \cos\frac{\theta}{2} & -\sin\frac{\theta}{2} \\ \sin\frac{\theta}{2} & \cos\frac{\theta}{2} \end{array}\right).
\end{equation}
To obtain $V_1$  we need only to swap the columns, i.e.
\begin{equation}
V_1 = \operatorname{RY}\left(\frac{3}{4} \pi \right) \operatorname{X}.
\end{equation}
Finally, the optimal probability of correct discrimination is equal to
\begin{equation}
p_{\text{succ}}(\PP_{U}, \PP_{\Id}) = \frac{1}{2} + \frac{\sqrt{2}}{4}.
\end{equation}
We will now demonstrate how to implement this theoretical scheme in PyQBench. For this example we
will use the Qiskit Aer simulator \cite{aer}. First, we import the necessary functions and classes from
PyQBench and Qiskit. We also import \texttt{numpy} for the definition of \texttt{np.pi} constant and
the exponential function. The exact purpose of the imported functions will be described at the point
of their usage.
\begin{lstlisting}[language=Python, caption=Imports needed for running benchmarking example]
import numpy as np
from qiskit import QuantumCircuit, Aer
from qbench.schemes.postselection import benchmark_using_postselection
from qbench.schemes.direct_sum import benchmark_using_direct_sum
\end{lstlisting}
To implement the discrimination scheme in PyQBench, we need to define all the necessary components
as Qiskit instructions. We can do so by constructing a circuit object acting on qubits 0 and 1 and
then converting them using \texttt{to\_instruction()} method.
\begin{lstlisting}[language=Python, caption= Defining components for Hadamard experiment]
def state_prep():
	circuit = QuantumCircuit(2)
	circuit.h(0)
	circuit.cnot(0, 1)
	return circuit.to_instruction()

>>>>

\subsection{Using user-defined measurement with \texttt{qbench} package}
In this example, we will demonstrate how \texttt{qbench} package can be used with user--defined
measurement. For this purpose, we will use $U = H$ (the Hadamard gate). The detailed calculations
that lead to the particular form of the discriminator and final measurements can be found in
\ref{app:hadamard}.
\noindent
The explicit formula for discriminator in this example reads:
\begin{equation}
\ket{\psi_0} = \frac{1}{\sqrt{2}} (\ket{00} + \ket{11}),
\end{equation}
with final measurements being equal to
\begin{equation}
V_0 =
\left(\begin{array}{cc} \alpha & -\beta\\ \beta & \alpha \end{array}\right),
\end{equation}
and \begin{equation}
V_1 =
\left(\begin{array}{cc} -\beta & \alpha \\ \alpha & \beta \end{array}\right),
\end{equation}
where \begin{equation}
\alpha = \frac{\sqrt{2 - \sqrt{2}}}{2} = \cos\left( \frac{3}{8} \pi \right),
\end{equation}
\begin{equation}
\beta  = \frac{\sqrt{2  + \sqrt{2}}}{2} = \sin\left( \frac{3}{8} \pi \right).
\end{equation}
To use the above benchmarking scheme in PyQBench, we first need to construct circuits that can be executed by actual hardware. To this end, we need to represent each of the unitaries as a sequence of standard gates, keeping in mind that quantum circuits start execution from the  $\ket{00}$ state. The circuit taking $\ket{00}$ to the Bell state $\ket{\psi_0}$  comprises the Hadamard gate followed by CNOT gate on both qubits (see Fig.~\ref{fig:discriminator}).
\begin{figure}[h!]
	\centering
	\includegraphics[scale=1.7]{pics/discriminator}
	\caption{Decomposition of the Bell state $\ket{\psi_0}$. }
	\label{fig:discriminator}
\end{figure}
For $V_0$ and $V_1$ observe that $V_0 = \operatorname{RY}\left(\frac{3}{4} \pi \right) $, where $\operatorname{RY}$ is rotation gate around the $Y$ axis defined by
\begin{equation}
\operatorname{RY}(\theta) =
\left(\begin{array}{cc} \cos\frac{\theta}{2} & -\sin\frac{\theta}{2} \\ \sin\frac{\theta}{2} & \cos\frac{\theta}{2} \end{array}\right).
\end{equation}
To obtain $V_1$  we need only to swap the columns, i.e.
\begin{equation}
V_1 = \operatorname{RY}\left(\frac{3}{4} \pi \right) \operatorname{X}.
\end{equation}
Finally, the optimal probability of correct discrimination is equal to
\begin{equation}
p_{\text{succ}}(\PP_{U}, \PP_{\Id}) = \frac{1}{2} + \frac{\sqrt{2}}{4}.
\end{equation}
We will now demonstrate how to implement this theoretical scheme in PyQBench. For this example we
will use the Qiskit Aer simulator \cite{aer}. First, we import the necessary functions and classes from
PyQBench and Qiskit. We also import \texttt{numpy} for the definition of \texttt{np.pi} constant and
the exponential function. The exact purpose of the imported functions will be described at the point
of their usage.
\begin{lstlisting}[language=Python, caption=Imports needed for running benchmarking example]
import numpy as np
from qiskit import QuantumCircuit, Aer
from qbench.schemes.postselection import benchmark_using_postselection
from qbench.schemes.direct_sum import benchmark_using_direct_sum
\end{lstlisting}
To implement the discrimination scheme in PyQBench, we need to define all the necessary components
as Qiskit instructions. We can do so by constructing a circuit object acting on qubits 0 and 1 and
then converting them using \texttt{to\_instruction()} method.
\begin{lstlisting}[language=Python, caption= Defining components for Hadamard experiment]
def state_prep():
	circuit = QuantumCircuit(2)
	circuit.h(0)
	circuit.cnot(0, 1)
	return circuit.to_instruction()
